---
title: "Spotify Exploration"
author: "Julia Koh, Dana Korssjoen, Heeju Son, and Allison Wick"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(spotifyr)
source("api_keys.R")
source("scripts/table.R")
source("scripts/summary.R")

```

## Introduction
Our problem domain analyzes audio features and music genres and answers the questions: 

- How do audio features vary across genres?
- How do audio features vary across popular songs in different countries?

The purpose of the project is to explore the differences of audio features between genres and across different countries. Using data from the audio streaming platform, Spotify, we have created a bar chart, histogram, and map of how these differences can be visualized. The source of our data comes from the `spotifyr` package.

## Summary

```{r summary_info, echo = FALSE}

```

## Table

The following table shows the average danceability, energy, valence, acousticness, and tempo based on **music genre**. The table is ordered by which genre has the highest average _popularity ratings_ on a scale from 0-100. `r popular` music is highest on the charts. `r tempo` music has the fastest average tempo. `r valence` music has the highest average valence. `r energy` music has the highest average energy. `r acoustic` music is the most acoustic on average. `r dance` music has the highest average danceability.

```{r summary_table, echo = FALSE}
summary_table <- get_table_info(songs_df)
kable(summary_table)
```

## Chart 1

```{r chart_one, echo = FALSE}

```

## Chart 2

```{r chart_two, echo = FALSE}

```

##Chart 3

```{r chart_three, echo = FALSE}

```
